### 2024-07-31 COLT/colt.tex

\documentclass[anon,12pt]{colt2024}

\usepackage{enumitem}
\usepackage{multicol}

\def\R{{\mathbb{R}}}
\def\pr{{\rm Pr}}
\def\E{{\mathbb E}}
\def\X{{\mathcal X}}
\def\Y{{\mathcal Y}}
\def\H{{\mathcal H}}
\def\G{{\mathcal G}}
\def\B{{\mathcal B}}
\def\yh{{\widehat{y}}}
\def\bias{{\rm bias}}
\def\supp{{\rm supp}}
\def\dist{{\rm dist}}
\def\sign{{\rm sign}}
\def\vol{{\rm vol}}
\def\PL{{\mbox{\rm PL}}}

\newtheorem{thm}{Theorem}
%\newtheorem{lemma}[thm]{Lemma}
\newtheorem{cor}[thm]{Corollary}
%\newtheorem{claim}[thm]{Claim}
\newtheorem{defn}[thm]{Definition}
\newtheorem{assump}{Assumption}
%\newenvironment{proof}{\noindent {\sc Proof:}}{$\Box$ \medskip}

\newcommand{\comment}[2]{ {\bf #1 :} #2}
\newcommand{\yoav}[1]{\comment{Yoav}{\em #1}}
\newcommand{\sanjoy}[1]{\comment{blue}{Sanjoy}{#1}}

\DeclareMathOperator*{\argmax}{arg\,max}

%\setlength{\leftmargini}{0.5cm}
%\setlength{\leftmarginii}{0.25cm}
%\setlength{\leftmarginiii}{0.25cm}

\title[Active learning using region-based sampling]{Active learning using region-based sampling}

\usepackage{times}

\coltauthor{
\Name{Sanjoy Dasgupta} \Email{dasgupta@eng.ucsd.edu}\and
\Name{Yoav Freund} \Email{yfreund@eng.ucsd.edu}\\
\addr University of California, San Diego}

\begin{document}

\maketitle

\begin{abstract}%
We present a general-purpose active learning scheme that maintains a collection of neighborhoods of different sizes and uses label queries to identify those with a strong bias towards one particular label. When two such neighborhoods intersect and have different labels, the region of overlap is treated as a ``known unknown'' and is targeted in future active queries. We give label complexity bounds for this method that do not rely on assumptions about the data, and we instantiate them in several cases of interest.
\end{abstract}

\begin{keywords}
Active learning, nonparametric estimation.
\end{keywords}

\section{Introduction}

In \emph{active learning}, the starting point is a data set whose labels are hidden but can be obtained individually for a price. The goal is to label the data set, or to find a good classifier, at low cost.

We consider a formulation in which we have an arbitrary collection of points $X = \{x_1, \ldots, x_n\}$ from some space $\X$. We can request the label of any of these points $x$, in which case we get a value $y \in \{-1, +1\}$ with conditional distribution
$ \eta(x) = \E[y | x] = \pr(y=1|x) - \pr(y=-1|x) $
for some unknown function $\eta: \X \to [-1,1]$. The \emph{Bayes-optimal} label for $x$, which we shall denote $g^*(x)$, is $-1$ if $\eta(x) < 0$ and $+1$ if $\eta(x) > 0$; either label can be used if $\eta(x) = 0$.
%$$ g^*(x) = \left\{
%\begin{array}{cl}
%-1 & \mbox{if $\eta(x) < 0$} \\
%+1 & \mbox{if $\eta(x) > 0$} \\
%\mbox{either} & \mbox{if $\eta(x)= 0$}
%\end{array} \right.
%$$
We wish to find the Bayes-optimal labels for the given data points $X$.

More precisely, at the outset we have the data $X$ and parameters $0 < \gamma, \delta < 1$. We want a procedure that chooses the next point, or batch of points, to query. This procedure will be applied repeatedly and can be stopped at any time, whereupon labels $\yh(x)$ must be provided for all $x \in X$, including those that were queried. Ideally, these will be the Bayes-optimal labels $g^*(x)$; however, we will only be judged on points with non-negligible bias, $|\eta(x)| \geq \gamma$. The number of mistakes is:
$$ \sum_{x \in X} {\bf 1} \big(\yh(x) \neq g^*(x),\ |\eta(x)| \geq \gamma \big) .$$
The overall procedure is allowed to fail with probability $\delta$, to account for sampling error.

\subsection{Approach and challenges}

In our setting there is no underlying assumption about $g^*$, for instance that it follows a linear model. Thus we adopt a nonparametric approach.

We begin with a large collection of \emph{neighborhoods}, $\B$, of varying sizes (Fig.~\ref{fig:challenges}(a)); we will refer to these as \emph{balls}. We use label queries to assign \emph{biases} $\yh(B) \in \{-1,0,+1\}$ to balls $B \in \B$, beginning with the largest balls. A bias of $+1$ means that the average $\eta(x)$ in $B$ is significantly positive; bias $-1$ means it is significantly negative; and bias $0$ means the average $\eta$ is close to 0. As time goes on, we get biases for progressively smaller balls, as needed. The final label assigned to any point $x$ is based on the biases of balls containing $x$.

\begin{figure}
\begin{center}
\includegraphics[width=5in]{ideas.pdf}
\end{center}
\caption{(a) A collection of balls $\B$. Ball $B$ has bias $\yh(B) \in \{-1,0,+1\}$. (b) If one ball has bias $+1$ and the other $-1$, points in their intersection are targeted for active querying. (c) Biases of progressively smaller balls containing $x$ might flip sign, leading to \emph{mind changes} about the label of $x$.}
\label{fig:challenges}
\end{figure}

%, me and to use label queries to estimate the signs associated with these regions. Ultimately we hope to cover $\X$ by a patchwork of neighborhoods with $\eta(x)$ values of uniform sign. Without strong smoothness conditions, these neighborhoods could be of very different sizes. The most beneficial for active learning are neighborhoods that contain a lot of $X$, but to get good coverage we might have to include smaller neighborhoods. To address this heterogeneity, we group the regions $\B$ by size. We begin by estimating the signs of the largest of them, and then move on to progressively smaller neighborhoods as the need arises. When the label budget runs out, the process is stopped, and provisional labels are assigned to the individual points in $X$.

There are three key challenges to be addressed. The first is deciding \emph{where to query}. Initial random querying tells us the biases of the largest balls, which gives some information about the decision boundary. In what parts of $\X$ might it be helpful to know biases of smaller balls, warranting further label queries? Our approach is simple: if a ball $B$ of bias $+1$ intersects a ball $B'$ of bias $-1$ then the intersection $B \cap B'$ is a ``known unknown'' and is targeted for active querying  (Fig.~\ref{fig:challenges}(b)).

This differs from the more conventional strategy of actively querying neighborhoods $B$ whose average $\eta$-value is close to zero. Such $B$ fall into two categories: (1) the $\eta$ values are close to zero throughout $B$, and (2) $B$ consists of two sub-regions, one strongly positive and the other strongly negative. In the first case, there is little merit in querying further. But in the second case, there is a lot to be gained. Our approach allows these two cases to be distinguished.

%To distinguish these two cases, we use a collection of neighborhoods that are \emph{overlapping}. For instance, we might take $\B$ to be \emph{all} balls in $\X$, which is effectively a finite collection once the given data points $X$ are taken into account. Case (2) can then be detected: we think of a point $x$ as being in the uncertainty region if it is contained in a neighborhood $B$ that is strongly positive as well as being in a neighborhood $B'$ that is strongly negative. Such points are ``known unknowns'', and these are the targets of our active querying.

A second challenge is that the sign of $\eta(\cdot)$ can change in an arbitrarily small neighborhood of any given point: the biases of balls containing $x$, say $B_1 \supset B_2 \supset \cdots$, might change sign a few times. At first, given only the bias of $B_1$, we might think $x$ has label $+1$. When we later estimate the bias of $B_2$, we could change our mind to $-1$. And then $+1$, and so on (Fig.~\ref{fig:challenges}(c)).

What this means is that we can \emph{never} be sure of having correctly determined the label of $x$. Thus in addition to focused (active) querying, we also do background sampling of the entire space to pick up on possible mind changes. For simplicity, we make one background query per focused query.

The third challenge is \emph{managing sampling} of overlapping balls. We are interested in detecting points, and thus neighborhoods, whose $\eta$-values are either $> \gamma$ or $<-\gamma$. This suggests querying $k \approx 1/\gamma^2$ points at random from each ball. Now, suppose we have queried this many points from ball $B$ and later want to query points from a different ball $B'$ that overlaps $B$. How can we reuse queries already made in $B \cap B'$?

\emph{Poisson sampling} provides a clean solution. Rather than choosing $k$ points at random from $B$, we pick each point in $B$ with probability (roughly) $k/|B \cap X|$, independently. We implement this by assigning each $x \in X$ a uniform-random value $T_x \in [0,1]$, at the outset. When sampling $B$, we choose to query $x$ if $T_x \leq k/|B \cap X|$. And when it comes time to sample a different $B'$ that also contains $x$, we choose $x$ if $T_x \leq k/|B' \cap X|$; if its label has already been obtained, we are able to reuse it. In this way, the random querying of overlapping balls is seamlessly managed.

These three challenges go beyond earlier work in active learning, which was able to avoid problems like mind-changes by making smoothness assumptions on $\eta$. By tackling all three of them, we are able to give a general-purpose active learning scheme.

\subsection{Results}

%Our algorithm does not require any assumption on the input distribution. In the worst case the number of queries is twice the number of queries of a passive algorithm. We show that under several natural conditions the query complexity of our algorithm is logarithmic in the error.

Section~\ref{sec:algorithm} presents the active learning algorithm and Section~\ref{sec:discrete} analyzes it in generality, taking $X$ to be an arbitrary point-set and allowing any $\eta$ function. We identify two \emph{critical levels} for any $x \in X$: two scales (of neighborhood sizes) that control how many queries are sufficient for $x$ to be correctly labeled (Theorem~\ref{thm:label-complexity}). We instantiate these bounds in canonical one-dimensional settings (Theorems~\ref{thm:oned-massart} and \ref{thm:oned-monotonic}) to get label complexities logarithmic in $|X|$.

Section~\ref{sec:continuous} considers the statistical setting where $X$ is drawn from an underlying distribution $\mu$ on $\X$. We give rates of convergence under distributional conditions (Theorem~\ref{thm:label-complexity-specific}), and instantiate them under common assumptions from learning theory and computational geometry.

\subsection{Related work}

Existing works on nonparametric active learning can mostly be grouped by overall principle: either (1) they seek to obtain the Bayes-optimal labels of a few well-positioned points, and then propagate these to the rest of the space \citep{DNZ15,H17,ASU18} or (2) they estimate the biases (positive or negative) of entire regions at a time \citep{DH08,M12}. In this paper, we follow the second strategy because the only reliable and general-purpose way to assess the bias of an individual point, $\mbox{sign}(\eta(x))$, is to query that point repeatedly; absent smoothness assumptions, the sign can change abruptly in an arbitrarily small neighborhood around $x$. On the other hand, the bias of a region $B$---the sign of the average $\eta$ value in $B$---is easy to determine given a reasonable number of points from that region.

Early results of \cite{CN08} established upper and lower bounds on label complexity in situations where the Bayes-optimal boundary is of a simple form: a single threshold for one-dimensional data or a ``smooth boundary fragment'' in higher dimension.

An algorithm for active learning based on hierarchical sampling was given by \cite{DH08} and was analyzed under smoothness conditions by \cite{KUB15}. The idea is to begin with a hierarchical clustering of $X$, and to then use queries to discover a pruning of this tree whose leaf-clusters are almost-pure in their labels. The method is not well-suited to situations with significant noise levels. Another approach using dyadic partitions was given by \cite{M12} and analyzed under commonly-used smoothness, margin, and density assumptions---namely, that $\eta$ is Holder-smooth, the fraction of points with $|\eta(x)| \leq t$ is some polynomial in $t$, and the marginal density is close to uniform---along with an additional smoothness requirement on $\eta$.

A different strategy using nearest neighbors was explored by \cite{ASU18}. Their idea was to choose an appropriate scale $s$, find an $s$-cover of $X$, estimate the Bayes-optimal label for each point in this cover by querying its neighbors, and then use these cleanly-labeled points for 1-nearest neighbor classification. A somewhat more general approach was given by \cite{H17} and studied under the usual smoothness, margin, and density conditions, with resulting rates of convergence comparable to those found by \cite{M12}.

Finally, \cite{DNZ15} suggested a graph-based method for active learning based on adaptively looking for the cut in the graph corresponding to the correct decision boundary. Their assumptions are based on properties of this cut and are not easily comparable with earlier work.

The idea of predicting with balls of multiple sizes is a central theme of \cite{BDFM19}, which predicts at a point $x$ by choosing a ball centered at $x$ that is as small as possible while having a clear bias towards one label. That work does not consider active learning, which demands considerable further innovation. For instance, to assess the uncertainty at a point $x$, we need to look at balls containing $x$, not just those centered at $x$ (Fig.~\ref{fig:challenges}(b)).

\section{The active learning algorithm}
\label{sec:algorithm}

Recall that we have a collection of points $X$ from a space $\X$. The label of any $x \in X$ can be requested and the value returned will be either $-1$ or $+1$, according to the conditional probability function $\eta(x) = \E[y|x]$. We wish to assign Bayes-optimal labels to all points in $X$ with $|\eta(x)| \geq \gamma$.

\subsection{A collection of sampling regions}

In our active learning algorithm, sampling is organized around a collection $\B$ of subsets of $\X$. These are the atomic sets on which we assess label bias and we call them \emph{balls} or \emph{neighborhoods}. There are no formal requirements on $\B$---for instance, the so-called balls can be of arbitrary shape---but the intention is that each $x \in X$ is contained in balls of multiple sizes, including one that is so small as to exclude the rest of $X$.

For any ball $B \in \B$, let $X_B = X \cap B$ be a shorthand for the data points that lie in it.
%\footnote{Since $X_B$ is used to assess the bias of $B$, we need to exclude points that cannot be considered random samples from $B$. For instance, consider these two alternatives: (1) $\B$ consists of a pre-defined set of balls. (2) $\B$ consists of balls defined by pairs of points in $X$, with each $x,x' \in X$ yielding the ball $B(x,\|x-x'\|)$. In the first case, all points $X \cap B$ are random draws from $B$. In the second case, we need to exclude the two points $x,x'$ that define the ball.} 
We group balls into levels by the number of points they contain. We put $B$ at {\it level} $\ell \geq 0$ if
\begin{equation}
\frac{n}{2^{\ell + 1}} \leq |X_B| < \frac{n}{2^\ell} .
\label{eq:sampling-level}
\end{equation}
Let $\B_\ell$ consist of all balls in $\B$ that are at level $\ell$. Thus $\B_0, \B_1, \ldots$ is a partition of $\B$, with $\B_0$ consisting of highly-populated balls and subsequent $\B_1, \B_2, \ldots$ consisting of successively smaller balls. We will use $\B_{\geq \ell}$ to denote all balls at levels $\ell$ or greater, and likewise $\B_{> \ell}$, $\B_{\leq \ell}$, and so on.

Balls in lower levels contain more points, and thus their biases (average $\eta$ values) are easier to estimate. Our sampling algorithm goes from easier to harder levels.

For any $x \in \X$, let $\B(x) \subset \B$ denote the collection of balls that contain $x$ and can thus be used in determining $x$'s label. We again partition these balls by sampling-level, so that $\B_\ell(x) = \B(x) \cap \B_\ell$. 
%We will also refer to $\B_\ell(x)$ as the {\em level $\ell$ neighborhood of $x$}.

\subsection{Estimating the biases of balls}

The bias of a ball $B \in \B$ is defined as the average $\eta$ value in it,
$$ \eta_X(B) = \mbox{average}\{\eta(x): x \in X_B \} .$$
We estimate these biases using label queries and assign each ball a \emph{qualitative} bias estimate,
$$ \yh(B) = 
\left\{
\begin{array}{cl}
+1 & \mbox{significant $+$ bias} \\
-1 & \mbox{significant $-$ bias} \\
0 & \mbox{no significant bias} \\
\bot & \mbox{not yet available}
\end{array}
\right.
$$

The option $\bot$ is used until enough points in $X_B$ have been queried: the required number is $k = O(1/\gamma^2)$ since $\gamma$ is the smallest bias that needs to be detected. Once this many labels are available, $\yh(B)$ is set to a value in $\{-1,0,+1\}$ and remains fixed thereafter. These bias estimates will with high probability be seen to satisfy the following guarantee.
\begin{defn}
For any $B \in \B$, bias estimate $\yh(B) \in \{+1,-1,0\}$ is \emph{$\gamma$-accurate} if:
\begin{itemize}
\item $\yh(B) = +1 \implies \eta_X(B) > 0$
\item $\yh(B) = -1 \implies \eta_X(B) < 0$
\item $\yh(B) = 0 \implies |\eta_X(B)| < \gamma$
\end{itemize}
\label{def:accurate-bias-estimate}
\end{defn}

\subsection{Assigning labels to points}

The label of any point is inferred from the biases of balls containing it.

Pick any $x \in X$ and any level $\ell$. Once qualitative bias estimates $\yh(B)$ are available for all balls at levels $\leq \ell$ that contain $x$ (that is, $\B_{\leq \ell}(x)$), we use these to infer a level-$\ell$ label for $x$. To begin with, the set of \emph{possible} labels for $x$ at level $\ell$, denoted $\PL_\ell(x) \subset \{-1,+1\}$, is defined thus:
\begin{itemize}[leftmargin=0.5cm]
\item $\PL_\ell(x)$ includes $+1$ if there exists a \emph{minimal} ball $B \in \B_{\leq \ell}(x)$ (with no other $B' \in \B_{\leq \ell}(x)$ strictly contained in it) for which $\yh(B) = +1$.
\item $\PL_\ell(x)$ includes $-1$ under a symmetrical condition.
\end{itemize}
This is made precise in Eqn~(\ref{eq:PL}) in Fig.~\ref{alg:main}. The label-estimate for $x$ at level $\ell$, denoted $\yh_\ell(x)$, is $+1$ if $\PL_\ell(x)=\{+1\}$, $-1$ if $\PL_\ell(x) = \{-1\}$, $0$ if $\PL_\ell(x) = \{\}$, and $!$ if $\PL_\ell(x) = \{-1,+1\}$ (Eqn~(\ref{eq:provisional-label})). 
%The interpretation of $\yh_\ell(x) \in \{+1,-1,0\}$ is given in Definition~\ref{def:accurate-bias-estimate}.

The label $\yh_\ell(x) = \, !$ indicates that at level $\ell$, there is conflicting evidence about $x$; it is a ``known unknown''~\citep{R11}. Our active learning algorithm makes its \emph{focused queries} at level $\ell+1$ in the vicinity of such points $x$, that is, in $\B_{\ell+1}(x)$.

% to be $-1$ if $\PL_\ell(x) = %\{-1\}$, or $+1$ if $\PL_\ell(x) = \{+1\}$, or $0$ if $\PL_\ell(x) = \{\}$, or $!$ %if $\PL_\ell(x) = \{-1,+1\}$, %and remains fixed thereafter. Points labeled $!$ are ``known unknowns'': they are %near the decision boundary and %need further investigation. Points labeled $0$ show no significant bias.

\subsection{The algorithm}

The active learning algorithm, shown in Figs~\ref{alg:main} and \ref{alg:sampling}, has two types of query. \emph{Background queries} are random draws from $X$ and correspond to passive learning. \emph{Focused queries} are made in the vicinity of ``uncertain'' points and correspond to active learning. The \emph{uncertainty region} $U_\ell$ at level $\ell \geq 1$ consists of points $x \in X$ on which the prediction from the previous level is ``known unknown'', $\yh_{\ell-1}(x) = \,!$. Focused queries are drawn at random from balls in $\B_\ell$ containing such points.

On each iteration of the main loop, (at most) one focused query is made as well as a background query. The focused query comes from the lowest-numbered uncertainty region $U_\ell$ that is nonempty.

Once these queries are made, bias estimates $\yh(B)$ are updated, along with label-sets $\PL_\ell(x)$, label-estimates $\yh_\ell(x)$, and finally the uncertainty regions $U_\ell$. Then the next iteration begins.

The querying process can be stopped at any time, whereupon labels are assigned as follows:
\begin{equation}
\yh(x) = 
\left\{
\begin{array}{cl}
\yh_\ell(x) & \mbox{largest $\ell$ with $\yh_\ell(x) \in \{-1,+1\}$, if such an $\ell$ exists} \\
0 & \mbox{``don't know'', otherwise}
\end{array}
\right.
\label{eq:final-label}
\end{equation}

\begin{figure}[t]
\framebox{
\begin{minipage}[t]{6in}
\begin{center}
\begin{minipage}[t]{6in}
\vspace{.05in}

Initialize uncertainty regions at all levels: $U_0 = X$ and $U_\ell = \emptyset$ for $\ell \geq 1$

\vspace{.1in}

Initialize labels at all levels to ``unavailable'': \ $\yh_\ell(x) = \bot$ for all $x$ and $\ell \geq 0$

\vspace{.1in}

Repeat:
\begin{itemize}[leftmargin=0.5cm]
\item If there is a level $\ell \geq 0$ such that $U_\ell \neq \emptyset$:
\begin{itemize}[leftmargin=0.25cm]
\item For the smallest such level $\ell'$, run {\tt Focused-query}($\ell', U_{\ell'}$) // Fig.~\ref{alg:sampling}
\end{itemize}
\item {\tt Background-query} // Fig.~\ref{alg:sampling}
\item Update labels:
\begin{itemize}[leftmargin=0.25cm]
\item Update bias-estimates $\yh(B)$ // Fig.~\ref{alg:bias-estimate}
\item For each $x \in X$ and level $\ell \geq 0$ for which all $\{\yh(B): B \in \B_\ell(x)\}$ are available:
\begin{align}
\PL_\ell(x) &= \{s \in \{-1, +1\}: \ \mbox{there exists $B \in \B_{\leq \ell}(x)$ s.t. $\yh(B) = s$ and} \notag \\ 
             &                       \hspace{1.5in}\mbox{no $B' \in \B_{\leq \ell}(x)$ has $X_{B'} \subsetneq X_B$.} \} \label{eq:PL} \\ 
\yh_\ell(x) &= 
\left\{
\begin{array}{cl}
+1 & \mbox{if $\PL_\ell(x) = \{+1\}$} \\
-1 &  \mbox{if $\PL_\ell(x) = \{-1\}$} \\
0 & \mbox{if $\PL_\ell(x) = \{\}$} \\
! & \mbox{if $\PL_\ell(x) = \{-1,+1\}$}
\end{array}
\right.
\label{eq:provisional-label}
\end{align}

\end{itemize}
\item Update uncertainty regions:
\begin{itemize}[leftmargin=0.25cm]
\item $U_0 = \{x \in X: \yh_0(x) = \bot\}$
\item For all levels $\ell \geq 1$: \ $U_\ell = \{x \in X: \yh_{\ell-1}(x) = \,!, \ \yh_\ell(x) = \bot\}$
\end{itemize}
\end{itemize}

\end{minipage}
\end{center}
\end{minipage}}
\caption{The active learning algorithm. Each iteration of the main loop makes (at most) one focused query and one background query.}
\label{alg:main}
\end{figure}

\subsection{Sampling}

The uncertainty region at level $\ell$ consists of the known-unknowns from level $\ell-1$, that is, points $x \in X$ with $\yh_{\ell-1}(x) = \, !$. This region is revealed to us piecemeal, a few points at a time. To get good label complexity we need to actively query the part of it we know.

We manage this through \emph{Poisson sampling} (Fig.~\ref{alg:sampling}). To infer the bias of a ball $B$, we need labels for $k = \tilde{O}(1/\gamma^2)$ random points in it. Balls at level $\ell$ have roughly $n/2^{\ell+1}$ points, so we need about $2^{\ell+1} k/n$ fraction of the ball to be queried. Instead, we query every point in the ball with probability $\tau_\ell = \min(2^{\ell+2}k/n, 1)$, \emph{independently}. Specifically, we query points $x$ in the ball with $T_x \leq \tau_\ell$, where each $x \in X$ is assigned a uniform-random value $T_x \in [0,1]$ at the outset of the algorithm.

Once a point is queried, it is never queried again.

\begin{figure}[t]
\framebox{
\begin{minipage}[t]{6in}

\vspace{.05in}
\emph{Initialization:}
\begin{itemize}[leftmargin=0.5cm]
\item Set $Q = \emptyset$ (points queried so far)
\item For each $x \in X$: choose $T_x \sim \mbox{uniform}([0,1])$
\end{itemize}

\vspace{.05in}
{\bf Background-query}

\begin{itemize}[leftmargin=0.5cm]
\item Query the next unlabeled point in $X \setminus Q$, ordered by increasing $T_x$ values, and add to $Q$
\end{itemize}

\vspace{.05in}
{\bf Focused-query}($\ell, U$)

\begin{itemize}[leftmargin=0.5cm]
\item Define querying region:
$$ S =  \bigcup_{x \in U} \bigcup_{B \in \B_\ell(x)} \{z \in X_B: T_z \leq \tau_\ell\} $$
\item Query the next unlabeled point in $S \setminus Q$, ordered by increasing $T_x$ values, and add to $Q$
\end{itemize}

\end{minipage}}
\caption{The two sampling procedures. Each $x \in X$ has an associated r.v. $T_x$ drawn uniformly from $[0,1]$. The smaller $T_x$, the earlier $x$ is likely to be queried. \emph{Background queries} are drawn at random from $X$. \emph{Focused queries} are from the uncertainty region at a given level $\ell$, and use level-based thresholds $\tau_\ell = \min(2^{\ell+2}k/n, 1)$, where $k = \tilde{O}(1/\gamma^2)$.} 
\label{alg:sampling}
\end{figure}

\begin{figure}
\framebox{
\begin{minipage}[t]{6in}
\vspace{.05in}
\begin{itemize}[leftmargin=0.25cm]
\item Initially $\yh(B) = \bot$
\item When all of $\{z \in X_B: T_z \leq \tau_\ell\}$ is queried, let $\widehat{\eta}(B)$ be the mean of these labels and set
$$ \yh(B)
= 
\left\{
\begin{array}{ll}
\sign(\widehat{\eta}(B)) & \mbox{if $|\widehat{\eta}(B)| \geq \gamma/2$} \\
0 & \mbox{otherwise}
\end{array}
\right.
$$
\end{itemize}
\end{minipage}}
\caption{The qualitative bias $\yh(B)$ of a ball $B \in \B_\ell$; see Fig.~\ref{alg:sampling} for $\tau_\ell$.}
\label{alg:bias-estimate}
\end{figure}

\section{Analysis: finite population setting}
\label{sec:discrete}

We now analyze the active learning procedure in a setting where $X \subset \X$ is an arbitrary set of $n$ points. We make no distributional assumption on the manner in which $X$ is generated.

\subsection{Accuracy of bias estimates}

Fix the set of balls $\B$ and let $0 < \delta < 1$ be a
predefined confidence parameter. We start with a uniform guarantee on the
bias estimates for all balls $B \in \B$.

In Fig.~\ref{alg:sampling}, we see that the points queried in a ball $B$ at level $\ell$ are $\{z \in X_B: T_z \leq \tau_\ell\}$, which from the definition of $\tau_\ell$ has size $O(k)$. We use these labels to estimate the bias of $B$. Since we need to detect biases of magnitude greater than $\gamma$, we set $k \propto 1/\gamma^2$. 
%This intuition is borne out by the following result, proved in the appendix. 
\begin{lemma}
Suppose that $k \geq (192/\gamma^2) \ln (4 |\B|/\delta)$. 
Let $\yh(B)$ be defined as in Fig.~\ref{alg:bias-estimate}. Then with probability $\geq 1-\delta$, all the $\yh(B)$, for $B \in \B$, are $\gamma$-accurate in the sense of Definition~\ref{def:accurate-bias-estimate}.
\label{lemma:accurate-bias-estimates}
\end{lemma}
Henceforth assume $\yh(B)$ is $\gamma$-accurate for all  $B \in \B$.

\subsection{Critical levels}

Recall that $\PL_\ell(x) \subset \{-1,+1\}$ denotes the possible labels for $x$ given information from balls at levels $\leq \ell$, and is used to assign a value $\yh_\ell(x) \in \{-1,+1,0,!\}$. Despite randomness in querying, it is possible to define two critical levels for each $x$: a level $L_1(x)$ by which $\PL_\ell(x)$ reliably contains the correct label of $x$, and a level $L_2(x)$ by which $\PL_\ell(x)$ reliably omits the wrong label.

\begin{defn}[Critical levels $L_1,L_2$]
Pick any $x \in X$ with $\eta(x) \neq 0$ and let $s(x) = \sign(\eta(x))$. Define $L_1(x)$ to be the smallest level $\ell$ such that there exists $B_o \in \B_\ell(x)$ for which
\begin{itemize}[leftmargin=0.5cm]
\item $s(x) \cdot \eta_X(B_o) \geq \gamma$, and
\item $s(x) \cdot \eta_X(B) \geq \gamma$ for any $B \in \B(x)$ with $X_{B} \subset X_{B_o}$.
\end{itemize}
Define $L_2(x)$ to be the smallest level $\ell$ such that:
\begin{itemize}[leftmargin=0.5cm]
\item For all $B \in \B_{\geq \ell}(x)$, we have $s(x) \cdot \eta_X(B) \geq 0$.
\item For any $B \in \B_{\leq \ell}(x)$ with $s(x) \cdot \eta_X(B) < 0$, there exists $B' \in \B_{\leq \ell}(x)$ with $X_{B'} \subset X_B$ and $s(x) \cdot \eta_X(B') \geq 0$.
\end{itemize}
Take $L_1(x)$ or $L_2(x)$ to be $\infty$ if no level meets the requirements, or if $\eta(x) = 0$. 
\label{defn:L12}
\end{defn}

Once $\ell$ reaches a critical level, $\PL_\ell(x)$ and thus $\yh_\ell(x)$ behave more predictably, as follows.
\begin{lemma}
Pick any $x \in X$ with $\eta(x) \neq 0$ and let $s(x) \in \{+1,-1\}$ denote its Bayes-optimal label. Then for any level $\ell$ and any time at which $\yh_\ell(x) \neq \bot$:
\begin{enumerate}
\item[(a)] If $\ell \geq L_1(x)$, then $s(x) \in \PL_\ell(x)$ and thus $\yh_\ell(x) \in \{s(x), !\}$. 
\item[(b)] If $\ell \geq L_2(x)$, then $-s(x) \not\in \PL_\ell(x)$ and thus $\yh_\ell(x) \in \{s(x), 0\}$.
\end{enumerate}
\label{lemma:boundary}
\end{lemma}

\subsection{Boundary sets and label complexity}

A common intuition about active learning is that successive queries gradually constrain the possible locations of the decision boundary. Let's consider the state of affairs when all balls at level $\leq \ell - 1$ have been sampled. The known-unknowns from this level are points $x$ with $\yh_{\ell-1}(x) = \, !$; by Lemma~\ref{lemma:boundary}(b), such points must have $L_2(x) \geq \ell$. Focused sampling at level $\ell$ is restricted to balls that contain these points. We will think of this region as the \emph{boundary set} at level $\ell$.
\begin{defn}[Boundary set $\bf \Delta_\ell$]
For any level $\ell$, define the \emph{boundary set} at level $\ell$ to be
\begin{equation}
\Delta_\ell 
= \bigcup_{x \in X: L_2(x) \geq \ell} \bigcup_{B \in \B_{\ell}(x)} X_B 
.
\label{eq:sampling-region}
\end{equation}
\end{defn}
%It can be shown that all focused samples at level $\ell$ must lie in this set.
\begin{lemma}
All focused samples at level $\ell$ lie in $\{z \in \Delta_\ell: T_z \leq \tau_\ell \}$.
\label{lemma:focused}
\end{lemma}

We can now give generic label complexity bounds in terms of $L_1$, $L_2$, and $\Delta_\ell$. These take two equivalent forms: a \emph{global} version that specifies what parts of $X$ are correctly labeled after $m$ queries and a \emph{local} version that specifies the number of queries after which a particular $x$ is correctly labeled. The global version (Theorem~\ref{thm:label-complexity-0}) is in the Appendix; here is the local version.

\begin{thm}[Local label complexity]
Take $k \geq (192/\gamma^2) \ln (4 |\B|/\delta)$. With probability at least $1-2\delta$, the following holds simultaneously for all $x \in X$. Let $L_1(x)$ and $L_2(x)$ be the critical levels for $x$, as in Definition~\ref{defn:L12}. If $L_2(x) \leq \lg (n/2k)$, let
\[
m_o(x) = 32k \cdot \max\bigg( 2^{L_1(x)}, \ \frac{1}{n} \sum_{\ell=L_1(x)+1}^{L_2(x)} |\Delta_\ell| \, 2^\ell\bigg) .
\]
Once the active learner has made $m_o(x)$ queries, $\yh(x)$ is fixed at the Bayes-optimal label $g^*(x)$.
\label{thm:label-complexity}
\end{thm}
Let's interpret this bound. For a specific $x \in X$, there is no telling how $\yh_\ell(x)$ will behave at levels $\ell < L_1(x)$. But for $\ell \geq L_1(x)$, either $\yh_\ell(x) = \, !$, in which case $x$ is in the uncertainty region, or the label of $x$ is correctly set once and for all, which occurs by level $L_2(x)$.

We will say that $x$ \emph{reaches level $\ell$} when bias estimates have been obtained for all balls in $\B_{\leq \ell}(x)$ and thus $\yh_\ell(x)$ is set. Recalling that the algorithm alternates between background and focused queries, we show that $m_o(x)/4$ \emph{background} queries suffice for $x$ to reach level $L_1(x)$ and an additional $m_o(x)/4$ \emph{focused} queries take it to level $L_2(x)$, by which time its label is correct.

In particular, $O(k \cdot 2^\ell)$ background queries give us the biases of all balls in $\B_{\leq \ell}$; this is the first term in $m_o(x)$. And there are $O(k \cdot 2^\ell \cdot |\Delta_\ell|/n)$ focused queries at any level $\ell$: the second term.

A footnote: Our algorithm queries points at most once. However, in some applications, a point $x$ can be queried repeatedly, each time producing an independent draw from $\eta(x)$. If \emph{repeat queries} are permissible, $O(1/\gamma^2)$ copies should be made of each point in $X$ before the algorithm is applied; and in this case, points with $|\eta(x)| \geq \gamma$ will all be correctly labeled, eventually.

\section{One-dimensional examples}

To apply Theorem~\ref{thm:label-complexity}, we need bounds on the critical levels $L_1(x)$ and $L_2(x)$ for each $x \in X$, and on the size of the sampling region $\Delta_\ell$ at each level $\ell$. We now derive these in some canonical settings.

\begin{figure}
\begin{center}
\includegraphics[width=2.5in]{oned-monotonic.pdf}
\hspace{.5in}
\includegraphics[width=2.5in]{oned-massart.pdf}
\end{center}
\caption{(a) The conditional probability function $\eta$ is monotonically increasing; $\eta(x) = -\gamma, 0, \gamma$ at $x = \lambda_L, \lambda, \lambda_R$, respectively. (b) On each subinterval, $\eta$ is either $> \gamma$ or $<-\gamma$.}
\label{fig:oned}
\end{figure}

\subsection{Example: monotonic $\eta$}

Suppose $X$ is an arbitrary set of $n$ points in $\X = [0,1]$ and is labeled according to a conditional probability function $\eta: \X \to [-1,1]$ that is continuous and strictly increasing (Fig.~\ref{fig:oned}(a)). Let $\lambda \in (0,1)$ be the point for which $\eta(\lambda) = 0$ and let $\lambda_L, \lambda_R$ be the points for which $\eta(\lambda_L) = -\gamma$ and $\eta(\lambda_R) = \gamma$. Thus we are not required to label points in the interval $(\lambda_L, \lambda_R)$.

%\begin{figure}
%\begin{center}
%\includegraphics[width=3in]{oned-monotonic.pdf}
%\end{center}
%\caption{The conditional probability function $\eta$ is monotonically increasing. The values $x = %\lambda_L, \lambda, \lambda_R$ have $\eta(x) = -\gamma, 0, \gamma$ respectively.}
%\label{fig:oned-monotonic}
%\end{figure}

Let $\B$ consist of all closed intervals. Although there are infinitely many of these, we need only consider those whose endpoints lie in $X$, so effectively $|\B| = O(n^2)$.

The critical levels are easy to bound in this setting. Let $n^- = |[0,\lambda_L] \cap X|$  be the number of points to the left of $\lambda_L$ and $n^+ = |[\lambda_R,1] \cap X|$ the number of points to the right of $\lambda_R$. Then 
%\begin{align*}
%L_1(x)
%&\leq
%\left\{
%\begin{array}{ll}
%\lg (n/n^+) & \mbox{if $x \geq \lambda_R$} \\
%\lg (n/n^-) & \mbox{if $x \leq \lambda_L$}
%\end{array}
%\right.
%\\
%L_2(x)
%&\leq
%\lg (n/r(x))
%\end{align*}
$$
L_1(x)
\leq
\left\{
\begin{array}{ll}
\lg (n/n^+) & \mbox{if $x \geq \lambda_R$} \\
\lg (n/n^-) & \mbox{if $x \leq \lambda_L$}
\end{array}
\right.
\mbox{\ \ \ \ and \ \ \ }
L_2(x)
\leq
\lg (n/r(x))
$$
where $r(x)$ is the number of data points between $x$ and $\lambda$, counting $x$ as well.
%\begin{itemize}
%\item $L_1(x) \leq \lg (n/n^+)$ if $x \geq \lambda_R$ and $L_1(x) \leq \lg (n/n^-)$ if %$x \leq \lambda_L$, and
%\item $L_2(x) \leq \lg (n/r(x))$, 
%\end{itemize}

A counting argument then shows that the boundary region shrinks exponentially as the level increases, $|\Delta_\ell| \leq 4n/2^\ell$, yielding the following label complexity. 
\begin{thm}
Define $r^+ = \min \{r(x): x \in X \cap [\lambda_R,1]\}$ and $r^- = \min \{r(x): x \in X \cap [0,\lambda_L]\}$. Pick any $0 < \delta < 1$. Suppose we run the algorithm of Fig.~\ref{alg:main} with $k = O((\log (n/\delta))/\gamma^2)$ and that $\min(r^+, r^-) \geq 2k$.  Take any 
$$ m \geq 64k \cdot \max \left( \frac{n}{\min(n^+,n^-)}, \ 2 \lg \frac{\min(n^+,n^-)}{\min(r^+,r^-)} \right) .$$
With probability $\geq 1-\delta$, after $m$ queries the algorithm correctly labels all $x \in X$ with $|\eta(x)| \geq \gamma$.
\label{thm:oned-monotonic}
\end{thm}
If $n^+, n^- = \Theta(n)$, that is, a constant fraction of the points lie to the left of $\lambda_L$ and to the right of $\lambda_R$, then the label complexity is $O(\log^2 n)$. If in addition $r^+, r^- = \Theta(n)$, that is, a constant fraction of the points lie between $\lambda_L, \lambda_R$ and the decision boundary, the label complexity is $O(\log n)$.

\subsection{Example: Massart noise}

In Section~\ref{sec:oned-massart}, we develop another one-dimensional example, in which $\X$ consists of several pieces, such that on each piece $\eta$ is either entirely above $\gamma$ or entirely below $-\gamma$ (Fig.~\ref{fig:oned}(b)). Once again, a logarithmic label complexity is obtained.

\section{Analysis: distributional setting}
\label{sec:continuous}

We now turn to a setting where the points $X$ are drawn from some distribution $\mu$ on a topological space $\X$. For the same algorithm, we bound label complexity in terms of properties of $\mu$ and $\eta$.

\subsection{A sufficiently-rich collection of neighborhoods}

Let $\B$ be a finite collection of subsets of $\X$; we spell out a basic property we need it to possess.

\begin{defn}
We say $\B$ is \emph{$k$-conforming} with respect to $X$ if 
\begin{itemize}
\item ({\bf Richness}) For any $x \in X$, there exists $B \in \B_0(x)$. Moreover, for any level $0 \leq \ell \leq \lg (n/k)$ and any $B \in \B_\ell(x)$, there exists $B' \in \B_{\ell+1}(x)$ such that $X_{B'} \subset X_B$.
\item ({\bf Fidelity}) $\mu(B)/2 \leq |X_B|/n \leq 2\mu(B)$ for all $B \in \B$ with $|X_B| \geq k$.
\end{itemize}
\label{defn:conformity}
\end{defn}
When $\B$ consists of Euclidean balls in $\R^d$, made finite by restricting to $X$, the first property is immediate and the second holds with high probability (Lemma~\ref{lemma:ball-size-bounds}). We also envisage scenarios where $\B$ is constructed using $X$, for instance from the cells of multiple spatial partition trees.

Henceforth assume $\B$ satisfies this property.  We ignore $B$ with $|X_B| < k$, thereby restricting attention to levels $0 \leq \ell < \lg (n/k)$.

\subsection{A probabilistic notion of distance}

A crucial step in the analysis is to introduce a distance function based on $\mu$: the distance from a point $x$ to a set $S$ is the probability mass of the smallest ball that contains $x$ and touches $S$.
\begin{defn}
For any $x \in \X$ and $S \subset \X$, define
$\dist(x,S) = \inf \{\mu(B): B \in \B(x), B \cap S \neq \emptyset\} .$
\label{defn:prob-dist}
\end{defn}
We will see that $L_2(x)$ can be bounded in terms of such distances. In what follows, for $s \in \{-1,0,+1\}$, take $\X^s$ to denote $\{x \in \X: \sign(\eta(x)) = s\}$.
\begin{lemma}
For any $x \in \X$, let $s(x) = \sign(\eta(x))$. If $\dist(x, \X^{-s(x)}) \geq 2k/n$, then
$L_2(x) \leq \bigg\lceil \lg \frac{1}{\dist(x, \X^{-s(x)})} \bigg\rceil + 1$.
\label{lemma:L2-bound}
\end{lemma}
We will also use probabilistic distance to define notions of boundary. Take the $p$-boundary to consist of points at distance $\leq p$ from points of the opposite label.
\begin{defn}
For any $p > 0$, define $\partial_p = \{x \in \X: \dist(x, \X^{-s(x)}) \leq p\}$.
\label{defn:boundary}
\end{defn}
This includes all $x$ with $\eta(x) = 0$. Next, the $(p,q)$-boundary consists of points at distance $\leq q$ from the $p$-boundary.
\begin{defn}
For any $p,q > 0$, define $\partial_{p,q} = \{x \in \X: \dist(x, \partial_p) \leq q\}$.
\label{defn:boundary2}
\end{defn}

We can bound $\Delta_\ell$, the query region at level $\ell$, in terms of this second-order boundary.
\begin{lemma}
For any $0 \leq \ell \leq \lg (2n/k)$, we have $\Delta_\ell \subset \partial_{4/2^\ell, 2/2^\ell}$.
\label{lemma:delta-continuous}
\end{lemma}
Theorem~\ref{thm:label-complexity} thus still holds with $n \cdot \mu(\partial_{4/2^\ell, 2/2^\ell})$ in place of $|\Delta_\ell|$; see Theorem~\ref{thm:label-complexity-dist} in the Appendix.

Using these notions, we are able to specialize the generic label complexity bound of Theorem~\ref{thm:label-complexity} to get rates under suitable margin conditions; this is given in Theorem~\ref{thm:label-complexity-specific} in the appendix. Here we just give a concrete scenario handled by the latter result; another such scenario, with smoothness and Tsybakov noise conditions, appears in the appendix.

\begin{figure}[t]
\begin{center}
\includegraphics[width=6in]{COLT/curvature.pdf}
\end{center}
\caption{{\bf A 2D example with bounded curvature:} {\em Left:} We have $\eta = -0.2$ for the orange/red part of the domain and $\eta=0.2$ for the light and dark blue. The orange and light blue indicate the boundary sets for negative and positive respectively. {\em Right:} examples of the $L_1,L_2$ balls for different points. The green $L_1$ circles are the largest monochrome circle that contains the point. The purple $L_2$ circles are the union of all circles of half the radius or smaller. All of these circles are monochrome.}
\label{fig:curvature}
\end{figure}

\subsection{Example: Curvature and Massart noise}
\label{sec:Massart}

By way of example, consider a case where $\X \subset \R^d$ and $\mu$ admits a density. We take $\B$ to consist of all open balls $B(x,r) = \{z: \|z-x\| < r\}$ centered in $\X$, with $\B(x)$ being balls that contain $x$. By VC arguments, the size of $\B$ is effectively $O(n^{d+1})$; in practice, this can be reduced to a low-order polynomial using core-sets~\citep{BC03}.

We make two key assumptions: distribution $\mu$ on $\X$ is within a multiplicative factor of uniform; and all of $\X$ has $|\eta(x)| \geq \gamma$ except for a $(d-1)$-dimensional boundary of finite curvature.
\begin{itemize}[leftmargin=0.5cm]
\item (Strong density condition)
There are constants $c_o, c_1 > 0$ such that for all balls $B \in \B$,
we have $c_o \vol(B) \leq \mu(B) \leq c_1 \vol(B) $, where $\vol(\cdot)$ is $d$-dimensional volume.
\item (Massart noise and boundary condition) $\X = \X^+ \cup \X^- \cup \X^0$, where: $\eta(x) \geq \gamma$ throughout $\X^+$; $\eta(x) \leq -\gamma$ throughout $\X^-$; and $\X^0$ separates (intersects any line between) $\X^+$ and $\X^-$ and is a $(d-1)$-dimensional Riemannian manifold of reach $r_o > 0$.
\end{itemize}
The \emph{reach} condition says that any point at distance $< r_o$ of $\X^0$ has a unique nearest neighbor on $\X^0$. It is a common notion of curvature~\citep{F59,NSW06} and implies that $\X^+$ (resp., $\X^-$) can be covered by open balls of radius $r_o$ that do not touch $\X^- \cup \X^0$ (resp., $\X^+ \cup \X^0$). A 2D example obeying these conditions appears in Figure~\ref{fig:curvature}.
\begin{thm}
Under the two conditions above, there are constants $c_2, c_3$ for which the following holds. Pick $0 < \delta < 1$ and take $k = O(((d \log n) + \log (1/\delta))/\gamma^2)$. If the algorithm of Fig.~\ref{alg:main} makes $c_2k \leq m \leq c_3 n^{(d-1)/d}$ queries, it assigns Bayes-optimal labels to all but $C (k/m)^{1/(d-1)} + (2/n) \log (4/\delta)$ fraction of $X$.
\label{thm:massart-dist}
\end{thm}
Here the constant $c_2$ includes a term $(1/r_o)^d$. The $\tilde{O}(1/m^{1/(d-1)})$ rate of this theorem is an improvement over the usual $1/m^{1/d}$ rate for random querying and was also observed earlier, in a more restricted setting, using specialized algorithms~\citep{CN08}.

\newpage

%\input{active-knn.bbl}
\bibliography{refs}
%\bibliographystyle{plain}

\newpage
\appendix

\section{Technicalities: discrete setting}

\subsection{Large deviation bounds for the discrete setting}

\begin{lemma}
Fix a confidence parameter $0 < \delta < 1$ and a positive integer $k \geq 6 \ln (4/\delta)$.

Let $x_1, \ldots, x_m$ be any collection of points. Suppose that the labels $Y_i \in \{-1,1\}$ of these points are independent, with $\E Y_i = \eta(x_i)$. Define
$$ \eta_o = \frac{1}{m} \left( \eta(x_1) + \cdots + \eta(x_m) \right) .$$
Now consider the following estimator $Z$ of this quantity:
\begin{itemize}
\item Each $x_i$ is chosen with probability $q > 0$, independently. Let $N$ be the number of selected points.
\item If $N > 0$, the labels $Y_i$ of the selected points are obtained, and $Z$ is their average.
\end{itemize}
If $qm \geq k + \sqrt{6k \ln (4/\delta)}$, with probability at least $1-\delta$, 
\begin{enumerate}
\item[(a)] $N \geq k$, and 
\item[(b)] $| Z - \eta_o| < \sqrt{(48/k) \ln (4/\delta)}$.
\end{enumerate}
\label{lemma:large-deviation-discrete}
\end{lemma}
\begin{proof}
Let's start with (a). Define $c = \sqrt{3 \ln (4/\delta)}$. We'll take $qm = k + \sqrt{6k \ln (4/\delta)} = k + c \sqrt{2k}$ since this is the worst case. By assumption, $k \geq 2c^2$ and thus $qm \leq 2k$.

Now, $N$ has a binomial($m,q$) distribution. By a multiplicative Chernoff bound, for $0 < \epsilon < 1$, we have
\begin{align*}
\pr(N \geq qm(1+\epsilon)) &\leq e^{-qm\epsilon^2/3} \\
\pr(N \leq qm(1-\epsilon)) &\leq e^{-qm\epsilon^2/2}
\end{align*}
Take $\epsilon = c/\sqrt{qm}$; by the lower bound on $k$, we have $\epsilon \leq 1/2$. Recalling the choice of $c$, we see that with probability at least $1-\delta/2$, we get 
$$(1-\epsilon) qm < N < (1+\epsilon) qm .$$
The lower bound implies $N > qm(1-\epsilon) = qm - c\sqrt{qm} = k + c \sqrt{2k} - c \sqrt{qm} \geq k$.

For (b), define $C_1, \ldots, C_m \in \{0,1\}$ as random variables indicating whether the corresponding points were selected; that is, $C_i = {\bf 1}(\mbox{$x_i$ was selected})$. The sum of the obtained labels is then $S = C_1 Y_1 + \cdots + C_m Y_m$. Notice that these $C_iY_i \in \{-1,0,1\}$ are independent with $\E[C_iY_i] = q \eta(x_i)$ and $\E[(C_iY_i)^2] = \E[C_i] = q$. Thus their sum $S$ has expectation
$$ \E [S] = \sum_{i=1}^m \E[C_i Y_i] = \sum_{i=1}^m q \eta(x_i) = qm \eta_o $$
and variance
$$ \mbox{var}(S) = \sum_{i=1}^m \mbox{var}(C_iY_i) \leq qm .$$
We can bound the concentration of $S$ around its expected value using Bernstein's inequality, by which
$$ \pr(|S - \E[S]| \geq t) \leq 2 \exp \left( - \frac{t^2}{2(\mbox{var}(S) + t/3)} \right) .$$
Using $t = \epsilon qm$, we then have that $|S - qm \eta_o| \leq \epsilon qm$ with probability at least $1-\delta/2$.

Combining with the high-probability bound on $N$ above, we get
$$ \frac{qm \eta_o - \epsilon qm}{qm(1+\epsilon)} < \frac{S}{N} < \frac{qm \eta_o + \epsilon qm}{qm(1-\epsilon)},$$
whereupon (recalling $Z = S/N$)
$$ 
\eta_o \left( \frac{1}{1+\epsilon} -1 \right) - \frac{\epsilon}{1+\epsilon} < Z - \eta_o < \eta_o \left( \frac{1}{1-\epsilon} - 1 \right) + \frac{\epsilon}{1-\epsilon} .$$
Since $|\eta_o| \leq 1$,
$$ |Z - \eta_o | < \max \left( \frac{2\epsilon}{1+\epsilon}, \frac{2\epsilon}{1-\epsilon} \right)
\leq 
4 \epsilon,$$
as claimed.  
\end{proof}

\subsection{Accuracy of bias estimates: Proof of Lemma~\ref{lemma:accurate-bias-estimates}}

We will use Lemma~\ref{lemma:large-deviation-discrete} to obtain a uniform guarantee on the bias estimates for all balls $B \in \B$.

Recall that each point $x \in X$ gets a label $Y \in \{-1,+1\}$ according to the distribution
$$ \eta(x) = \E[y | x].$$ 
For any $B \in \B$ with $X_B \neq \emptyset$, let $\eta_X(B)$ be the average $\eta$-value of the points in $B$, that is,
$$ \eta_X(B) = \frac{1}{|X_B|} \sum_{x \in X_B} \eta(x) .$$

For any $B \in \B_\ell$, define its \emph{query-set} to be $\Gamma(B) = \{z \in X_B: T_z \leq \tau_\ell\}$, where the sampling threshold $\tau_\ell$ for level $\ell$ is defined as:
\begin{equation}
\tau_\ell = \min \left( \frac{2^{\ell+2}k}{n}, \ 1 \right)
\end{equation}
for some $k$. We base our bias-estimate for $B$ on the labels of points in $\Gamma(B)$.

For what follows, define
\begin{equation*}
c = \sqrt{3 \ln \frac{4|\B|}{\delta}} .
\end{equation*}

\begin{lemma}
Suppose $k \geq 2c^2$. With probability at least $1-\delta$, the following is true for all $B \in \B$ with $|X_B| \geq k$:
\begin{enumerate}
\item[(a)] The query set $\Gamma(B)$ has size at least $k$.
\item[(b)] The average label on $\Gamma(B)$, call it $\widehat{\eta}(B)$, satisfies
$$ \left| \widehat{\eta}(B) - \eta_X(B) \right| \leq \frac{4c}{\sqrt{k}}.$$
\end{enumerate}
\label{lemma:large-deviation-bounds}
\end{lemma}
\begin{proof}
Pick $B \in \B$. There are two cases to consider.

Case 1: $|X_B| \geq 2k$. The choice of $\tau_\ell$ then ensures $|X_B| \tau_\ell \geq 2k$, so that $|\widehat{\eta}(B) - \eta_X(B)|$ can be bounded by applying Lemma~\ref{lemma:large-deviation-discrete} to the points $X_B$ with sampling probability $q = \tau_\ell$.

Case 2: $k \leq |X_B| < 2k$. In this case, $B$ lies at a level $\ell$ for which $\tau_\ell = 1$. Lemma~\ref{lemma:large-deviation-discrete} does not apply directly, but its conclusion still holds. In particular, the query-set $\Gamma(B)$ is all of $X_B$, and the same Bernstein bound from the proof of Lemma~\ref{lemma:large-deviation-discrete} can be again be applied.

To complete the proof, we take a union bound over all $B \in \B$.
\end{proof}

The following corollary of Lemma~\ref{lemma:large-deviation-bounds} is immediate.
\begin{cor}
Suppose that $k \geq (8c/\gamma)^2$. For each $B \in \B$, let $\widehat{\eta}(B)$ be the average label on the query-set $\Gamma(B)$, and define the bias-estimate $\yh(B)$ as follows:
$$ \yh(B)
= 
\left\{
\begin{array}{ll}
\sign(\widehat{\eta}(B)) & \mbox{if $|\widehat{\eta}(B)| \geq \gamma/2$} \\
0 & \mbox{otherwise}
\end{array}
\right.
$$
With probability $\geq 1-\delta$, all bias estimates $\yh(B)$, for $B \in \B$, are $\gamma$-accurate.
\label{cor:accurate-bias-estimates}
\end{cor}
\begin{proof}
By the choice of $k$, we have from Lemma~\ref{lemma:large-deviation-bounds} that 
$| \widehat{\eta}(B) - \eta_X(B) | < \gamma/2$
for all $B \in \B$.
\end{proof}

\subsection{Critical levels: Proof of Lemma~\ref{lemma:boundary}}

For part (a), note that some $B_o \in \B_{L_1(x)}(x)$ has significant bias (that is, bias $\geq \gamma$) towards the correct label $s(x)$, as does any ball $B \in \B(x)$ contained within it. For $\ell \geq L_1(x)$, the set $\{B \in \B_{\leq \ell}(x): X_{B} \subseteq X_{B_o}\}$ is nonempty (it contains $B_o$); pick any minimal ball $B$ within it. Then $\yh(B) = s(x)$ by the $\gamma$-accuracy of bias estimates (Lemma~\ref{lemma:accurate-bias-estimates}) and thus $s(x) \in \PL_\ell(x)$.

For (b), take any $\ell \geq L_2(x)$. Consider any $B \in \B_{\leq \ell}(x)$ for which $s(x) \cdot \eta_X(B) < 0$. By definition of $L_2(x)$, this $B$ must lie in $B_{< L_2(x)}(x)$, and moreover there must exist $B' \in \B_{\leq L_2(x)}(x) \subset \B_{\leq \ell}(x)$ with $X_{B'} \subset X_B$ and $s(x) \cdot \eta_X(B') \geq 0$. Thus, any minimal $B \in \B_{\leq \ell}(x)$ has bias $\geq 0$ towards the correct label $s(x)$, whereupon $\yh(B) \in \{0, s(x)\}$ by the $\gamma$-accuracy of bias estimates. Therefore, $-s(x) \not\in \PL_\ell(x)$.

\subsection{The region of focused sampling: Proof of Lemma~\ref{lemma:focused}}

Let $\overline{U}_\ell$ denote the set of all points that are ever (in any round of sampling) in the uncertainty set at level $\ell$. From Lemma~\ref{lemma:boundary}(b), we see that any $x$ with $L_2(x) < \ell$ has $\yh_{\ell-1}(x) \neq \ !$ and thus never makes it into the uncertainty set at level $\ell$. In short,
\begin{equation}
\overline{U}_{\ell} \subset \{x \in X: L_2(x) \geq \ell\}.
\label{eq:uncertainty}
\end{equation}

We see from the {\tt Focused-query} subroutine (Figure~\ref{alg:sampling}) that all focused samples at level $\ell$ lie in
$$
\bigcup_{x \in \overline{U}_\ell} \bigcup_{B \in \B_\ell(x)} \{z \in X_B: T_z \leq \tau_\ell \}
\ 
\subset 
\ 
\bigcup_{x \in X: L_2(x) \geq \ell} \bigcup_{B \in \B_\ell(x)} \{z \in X_B: T_z \leq \tau_\ell \}
\ 
=
\ 
\{z \in \Delta_\ell: T_z \leq \tau_\ell \}.
$$

\subsection{A generic label complexity bound}

We start by showing that various subsets of interest contain roughly the expected number of points at each level.
\begin{lemma}
With probability at least $1-2(\lg (n/k))e^{-k/3}$, the following hold for all levels $0 \leq \ell \leq \lg (n/2k)$.
\begin{enumerate}
\item[(a)] $|\{x \in X: T_x \leq \tau_\ell\}| < 2 n \tau_\ell$.
\item[(b)] If $\Delta_\ell \neq \emptyset$ then $|\{x \in \Delta_{\ell}: T_x \leq \tau_\ell\}| < 2 |\Delta_\ell| \tau_\ell$.
\end{enumerate}
\label{lemma:level-distribution}
\end{lemma}
\begin{proof} 
Pick any subset $S \subset X$ and let $m = |S|$. Then $|\{x \in S: T_x \leq \tau_\ell\}|$ has a $\mbox{binomial}(m, \tau_\ell)$ distribution with expectation $m \tau_\ell$. The probability that it is greater than or equal to twice its expected value is, by a multiplicative Chernoff bound, at most $e^{-m \tau_{\ell}/3}$, which is $\leq e^{-k/3}$ as long as $m \tau_\ell \geq k$.

Both parts follow from this principle; and we take a union bound over all $\lg (n/k)$ levels. For (b), we need to check that $|\Delta_\ell| \tau_\ell \geq k$. To see this, observe from the definition (\ref{eq:sampling-region}) of $\Delta_\ell$ that if it is non-empty, then it contains $X_B$ for at least one ball $B \in \B_\ell$, and every such ball has at least $n/2^{\ell+1}$ points. Combining this with the definition $\tau_\ell = \min(2^{\ell+2}k/n, 1)$ yields $|\Delta_\ell| \tau_\ell \geq k$.
\end{proof}

Theorem~\ref{thm:label-complexity} is a restated version of the following.
\begin{thm}
Suppose the active learning algorithm makes $0 < m \leq n$ queries. Then all points $x \in X$ with $L_1(x) \leq \ell_1$ and $L_2(x) \leq \ell_2$ will get Bayes-optimal labels $\yh(x) = g^*(x)$, where
$$ \ell_1 = \left\lfloor \lg \frac{m}{32k} \right\rfloor $$
and $\ell_2$ is the largest integer $\leq \lg (n/2k)$ such that
$$ \sum_{\ell = \ell_1 + 1}^{\ell_2} |\Delta_{\ell}| \, \tau_{\ell} \ < \ \frac{m}{8} .$$
\label{thm:label-complexity-0}
\end{thm}

\begin{proof}
Denote the first $m/2$ queries by {\it phase one} and the second $m/2$ by {\it phase two}. We will analyze the effect of background sampling in phase one and focused sampling in phase two. We start with the former.

Of the $m/2$ queries in phase one, at least $m/4$ will be background samples. Therefore the $m/4$ points with lowest $T_x$ values are guaranteed to be queried.

Now, for $\ell_1$ as defined, we have that $\tau_{\ell_1} \leq m/8n$ and thus from Lemma~\ref{lemma:level-distribution}(a) that at most $m/4$ points in $X$ satisfy $T_x \leq \tau_{\ell_1}$. Therefore all such points are queried in phase one, and all label-estimates $\{\yh_{\ell_1}(x): x \in X\}$ are set.

It then follows from Lemma~\ref{lemma:boundary}(a) that the following hold for any $x \in X$ with $L_1(x) \leq \ell_1$:
\begin{enumerate}
\item[(a)] By the end of phase one, $\yh_{\ell_1}(x) \in \{g^*(x), !\}$.
\item[(b)] For any $\ell > \ell_1$, when $\yh_\ell(x)$ becomes available, it lies in $\{g^*(x), !\}$.
\item[(c)] If $x$ ever leaves the combined uncertainty region $U = \cup_\ell U_\ell$ during phase two, then its final label as defined in (\ref{eq:final-label}) is henceforth always $\yh(x) = g^*(x)$.
\end{enumerate}

Now let's move on to phase two. Let $A = \{x \in X: L_1(x) \leq \ell_1, L_2(x) \leq \ell_2\}$. We will show that every point in $A$ must leave the uncertainty region $U$ at some time during phase two. From (c), we can conclude that all these points have their final labels set correctly, once and for all.

We will break the argument into two cases.

{\it Case 1:} Fewer than $m/4$ focused queries are made in phase two. This can only happen if some round of sampling has an empty uncertainty set, meaning that all of $A$ has left $U$ at that point.

{\it Case 2:} A full $m/4$ focused queries are made in phase two. By the analysis of phase one, none of these queries can be at level $\leq \ell_1$ and by Lemma~\ref{lemma:focused}, the total number of possible focused queries at levels $\ell_1+1$ through $\ell_2$ inclusive is at most
$$\sum_{\ell=\ell_1+1}^{\ell_2} |\{z \in \Delta_{\ell}: T_z \leq \tau_\ell \}| 
\ \leq \ 
\sum_{\ell=\ell_1+1}^{\ell_2} 2 |\Delta_{\ell}| \tau_\ell
\ < \ 
\frac{m}{4} ,
$$
where the first inequality is from Lemma~\ref{lemma:level-distribution}(b). Thus at least one query in phase two must be at level $\ell_2 +1$. When this query is made, every $U_\ell$ with $\ell \leq \ell_2$ must be empty and thus all of $A$ must have left the uncertainty region; recall from (\ref{eq:uncertainty}) that no $x \in A$ can be part of $U_\ell$ for $\ell > \ell_2 \geq L_2(x)$.

Thus every $x \in A$ must leave the uncertainty region at some point in phase two, and their final labels are subsequently set correctly.
\end{proof}

\subsection{One-dimensional monotonic $\eta$: Proof of Theorem~\ref{thm:oned-monotonic}}

We begin by bounding the critical levels $L_1$ and $L_2$ for points in $X$.
\begin{lemma}
Pick any $x \in [0, 1]$.
\begin{enumerate}
\item[(a)] Define $n^- = |[0,\lambda_L] \cap X|$ and $n^+ = |[\lambda_R,1] \cap X|$. Then
$$
L_1(x)
\leq
\left\{
\begin{array}{ll}
\lg (n/n^+) & \mbox{if $x \geq \lambda_R$} \\
\lg (n/n^-) & \mbox{if $x \leq \lambda_L$}
\end{array}
\right.
$$
\item[(b)] Let $r(x)$ be the number of points between $x$ and the boundary point $\lambda$, counting $x$ itself. That is, $r(x) = |[x,\lambda) \cap X|$ if $x < \lambda$, or $|(\lambda, x] \cap X|$ if $x > \lambda$. Then $L_2(x) \leq \lg (n/r(x))$.
\end{enumerate}
\label{lemma:oned-monotonic-L}
\end{lemma}
\begin{proof}
For (a), take any $x \geq \lambda_R$ (the other case is similar). The interval $B = [\lambda,1]$ lies in $\B_\ell(x)$ for $\ell = \lceil \lg (n/n^+) - 1 \rceil$ and has $\eta_X(B) \geq \gamma$. Furthermore, any $B' \subset B$ also has $\eta_X(B') \geq \gamma$.

For (b), take $x \geq \lambda_R$ and $\ell \geq \lg (n/r(x))$. Any $B \in \B_\ell(x)$ contains $< n/2^\ell \leq r(x)$ points and thus cannot possibly extend to the other side of the boundary. It follows that every $B \in \B_{\geq \ell}(x)$ has $\eta_X(B) \geq 0$. Moreover, any interval $B'$ that does extend to the other side of the boundary contains $B' \in \B_{\ell}(x)$ that is entirely on the same side as $x$.
\end{proof}

We can now bound the size of the query region at each level and find that it shrinks exponentially with $\ell$.
\begin{lemma}
For any $\ell \geq 0$, let $\Delta_\ell$ denote the focused querying region at level $\ell$, as defined in (\ref{eq:sampling-region}). Then $|\Delta_\ell| \leq 4n/2^\ell$.
\label{lemma:oned-monotonic-query-region}
\end{lemma}

\begin{proof}
We have
\begin{align*}
\Delta_\ell 
&= \bigcup_{x \in X: L_2(x) \geq \ell} \bigcup_{B \in \B_{\ell}(x)} (B \cap X) \\
&\subset \bigcup_{x \in X: r(x) \leq n/2^\ell} \bigcup_{B \ni x: |B \cap X| < n/2^\ell} (B \cap X) 
.
\end{align*}
This includes at most $n/2^{\ell-1}$ points from $X$ on either side of $\lambda$. \end{proof}

We are now ready for the proof of Theorem~\ref{thm:oned-monotonic}.

Setting $k$ to $O((1/\gamma^2) \ln (n/\delta))$ satisfies the requirements of Theorem~\ref{thm:label-complexity}. Here we are using the fact that although $\B$ is infinite, we need only consider $O(n^2)$ distinct intervals since $|X| = n$.

First observe that for any $\ell_1 \leq \ell_2$, we have from Lemma~\ref{lemma:oned-monotonic-query-region} that
$$ \sum_{\ell=\ell_1+1}^{\ell_2} |\Delta_\ell| \tau_\ell 
\ \leq \ \sum_{\ell=\ell_1+1}^{\ell_2} \frac{4n}{2^\ell} \cdot \frac{2^{\ell+2} k}{n} 
\ = \ 16k(\ell_2-\ell_1).$$
Now, let's define
$$ \ell_1 = \left\lfloor \lg \frac{m}{32k} \right\rfloor, 
\ \ \ \ell_2 = \min \left( \ell_1 + \frac{m}{128k}, \ \ \lg \frac{n}{2k} \right).$$
Then for any $x \in [0, \lambda_L] \cup [\lambda_R,1]$, we have
$$ \ell_1 = \left\lfloor \lg \frac{m}{32k} \right\rfloor 
\ \geq \ 
\lg \frac{m}{64k}
\ \geq \ 
\lg \frac{n}{\min(n^+,n^-)}
\ \geq \ 
L_1(x)$$
and, if $\min(r^+, r^-) \geq 2k$,
\begin{align*}
\ell_2 
\ = \ \ell_1 + \frac{m}{128k} 
&\geq \lg \frac{m}{64k} + \lg \frac{\min(n^+,n^-)}{\min(r^+,r^-)} \\
&\geq \ \lg \frac{m}{64k} + \lg \frac{64kn/m}{\min(r^+,r^-)} 
\ = \ \lg \frac{n}{\min(r^+,r^-)}
\ \geq \  
L_2(x).
\end{align*}
We get the algorithmic guarantee by applying Theorem~\ref{thm:label-complexity}.

\subsection{One-dimensional data with Massart noise}
\label{sec:oned-massart}

We now turn to another one-dimensional setting. Once again, $X$ consists of $n$ arbitrarily-placed points in $\X = [0,1]$. This time, however, they are labeled according to a conditional probability function $\eta: \X \to [-1,1]$ that satisfies the Massart noise condition:
\begin{itemize}
\item There are $p$ disjoint open intervals $I_1, \ldots, I_p$, such that $\X$ is (the closure of) their union, and
\item for each $j$, either $\eta(x) > \gamma$ for all $x \in I_j$ or $\eta(x) < -\gamma$ for all $x \in I_j$. In the first case, we write $s(I_j) = +1$ and in the second case, $s(I_j) = -1$.
\end{itemize}
Here $0 < \gamma < 1$ is some constant. See Figure~\ref{fig:oned}(b) for an illustrative example. For concreteness, the intervals $I_j$ can be written in the form $(\lambda_{j-1}, \lambda_j)$, where
$0 = \lambda_0 < \lambda_1 < \cdots < \lambda_{p-1} < \lambda_p = 1 .$
Here $\lambda_1, \ldots, \lambda_{p-1}$ are the \emph{boundary points} between intervals.

We will take $\B$ to consist of all open intervals of $[0,1]$, with $\B(x)$ denoting intervals that contain point $x$.

We begin with bounds on the $L_1$ and $L_2$ levels for each point.
\begin{lemma}
Pick any $x \in \X$; suppose $x \in I_j$.
\begin{enumerate}
\item[(a)] Let $n_j = |X \cap I_j|$. Then $L_1(x) \leq \lg (n/n_j)$.
\item[(b)] Let $r(x)$ be the minimum number of data points that lie between $x$ and a boundary point, counting $x$ as well; this is either the number of points in the left-interval $(\lambda_{j-1},x]$ (if $j > 1$) or the right-interval $[x, \lambda_j)$ (if $j < p$), whichever is smaller. Then $L_2(x) \leq \lg (n/r(x))$.
\end{enumerate}
\label{lemma:oned-massart-L}
\end{lemma}
\begin{proof}
For (a), notice first that $I_j \in \B_{\ell}(x)$ for $\ell = \lceil (\lg n/n_j) - 1 \rceil$. Moreover, $s(I_j) \cdot \eta_X(I_j) > \gamma$. Thus $I_j$ belongs to $\B_\ell(x)$ and is strongly biased towards the correct label. This strong bias also holds for any subset of $I_j$.

For (b), consider any $\ell \geq \lg (n/r(x))$. Any $B \in \B(x)$ with $\mbox{sign}(\eta_X(B)) \neq s(I_j)$ must contain either the entire left-interval $(\lambda_{j-1},x]$ or the entire right-interval $[x, \lambda_j)$, and thus has at least $r(x)$ points, which means that it is too large to be in $\B_\ell$. Thus all intervals $B \in \B_{\geq \ell}(x)$ have $s(I_j) \cdot \eta_X(B) > 0$. Also, for any interval $B \in \B_{<\ell}(x)$ there is some $B' \in \B_{\ell}(x)$ that is strictly contained within it.
\end{proof}

With $L_1(x)$ and $L_2(x)$ under control, it is easy to bound the size of the focused query region $\Delta_\ell$ at each level.
\begin{lemma}
For any $\ell \geq 0$, let $\Delta_\ell$ denote the focused querying region at level $\ell$, as defined in (\ref{eq:sampling-region}). Then $|\Delta_\ell| \leq (p-1)n/2^{\ell-2}$.
\label{lemma:oned-massart-query-region}
\end{lemma}

\begin{proof}
We have
\begin{align*}
\Delta_\ell 
&= \bigcup_{x \in X: L_2(x) \geq \ell} \bigcup_{B \in \B_{\ell}(x)} (B \cap X) \\
&\subset \bigcup_{x \in X: r(x) \leq n/2^\ell} \bigcup_{B \ni x: |B \cap X| < n/2^\ell} (B \cap X) 
.
\end{align*}
This includes at most $n/2^{\ell-1}$ points from $X$ on either side of each boundary point $\lambda_j$.
\end{proof}

Notice that $|\Delta_\ell|$ shrinks exponentially with $\ell$. With $L_1$, $L_2$, and $|\Delta_\ell|$ values in place, Theorem~\ref{thm:label-complexity} can be applied directly to give the following label complexity bound.

\begin{thm}
Pick any $0 < \epsilon, \delta < 1$. Suppose we run the algorithm of Figure~\ref{alg:main} with $k = O((1/\gamma^2) \ln (n/\delta))$. With probability at least $1-\delta$, after making
$$ O \left( \frac{p-1}{\gamma^2} \ln \frac{p-1}{\epsilon} \ln \frac{n}{\delta} \right) $$
queries, the algorithm will assign the correct label $g^*(x)$ to at least $1-\epsilon$ fraction of $X$, except possibly the $2k$ points of either side of each boundary point.
\label{thm:oned-massart}
\end{thm}

\begin{proof}
Setting $k$ to $O((1/\gamma^2) \ln (n/\delta))$ satisfies the requirements of Theorem~\ref{thm:label-complexity}. Here we are using the fact that although $\B$ is infinite, we need only consider $O(n^2)$ distinct intervals since $|X| = n$.

Next, using Lemma~\ref{lemma:oned-massart-query-region}, we have that for any integers $0 \leq \ell_1 < \ell_2$,
$$ \sum_{\ell = \ell_1+1}^{\ell_2} |\Delta_\ell| \tau_\ell
\ \leq \ 
\sum_{\ell = \ell_1+1}^{\ell_2} \frac{(p-1) n}{2^{\ell-2}} \cdot \frac{2^{\ell+2}k}{n}
\ = \ 
16(p-1)k (\ell_2 - \ell_1).
$$
We can then apply Theorem~\ref{thm:label-complexity} to conclude that $m$ query points are enough to correctly classify all $x \in X$ with $L_1(x) \leq \ell_1$ and $L_2(x) \leq \ell_2$, for
\begin{align*}
\ell_1
&= 
\left\lfloor \lg \frac{m}{32k} \right\rfloor \\
\ell_2
&=
\min \left( \ell_1 + \frac{m}{128 k(p-1)}, \ \lg \frac{n}{2k} \right)
\end{align*}
Using Lemma~\ref{lemma:oned-massart-L}, we have that in every target interval $I_j$ with $n_j/n = \Omega(k/m)$, all but $n \cdot 2^{-\ell_2-1}$ points will be correctly classified. For large enough $m$, this means that the fraction of misclassified or unclassified points in $X$ will be at most $\epsilon$ after $O(k(p-1) \log ((p-1)/\epsilon))$ queries, apart from the $2k$ points nearest the boundaries, which will remain unclassified.
\end{proof}

The $2k$ points nearest each boundary cannot be labeled by our algorithm with any certainty because they lie in intervals with a strongly positive bias as well as in intervals with a strongly negative bias. This qualification would be removed if were allowed to make multiple queries to each point, because in that case we would include $O(k)$ copies of each point, as explained earlier.

\section{Technicalities: distributional setting}

In the distributional setting, $X$ is drawn i.i.d.\ from a distribution $\mu$ on a topological space $\X$. We have a collection of regions $\B$.

\subsection{Sampling level and probability mass}

We start by showing how the fidelity of $\B$ (Definition~\ref{defn:conformity}) can be established.
\begin{lemma}
With probability at least $1-\delta$,
\begin{enumerate}
\item[(a)] For all $B \in \B$ with $n \mu(B) \geq 12 \ln (2|\B|/\delta)$, we have
$$ \frac{\mu(B)}{2} \leq \frac{|X_B|}{n} \leq 2 \mu(B) .$$
\item[(b)] For all $B \in \B$ with $n \mu(B) < 12 \ln (2|\B|/\delta)$, we have $|X_B| < 24 \ln (2 |\B|/\delta)$.
\end{enumerate}
\label{lemma:ball-size-bounds}
\end{lemma}
\begin{proof}
It is an immediate consequence of the multiplicative Chernoff bound that with probability at least $1-\delta$, for all $B \in \B$,
$$ \frac{|X_B|}{n} = \mu(B) \left(1 \pm \sqrt{\frac{3}{n \mu(B)} \ln \frac{2|\B|}{\delta}} \right) .$$
\end{proof}

In particular, taking $k > 24 \ln (2 |\B|/\delta)$ ensures that condition (a) hold will hold for all $B \in \B$ with $|X_B| \geq k$.

Next, we will see that balls of probability mass $p$ belong to level $\ell \approx \lg (1/p)$.
\begin{lemma}
For any level $\ell \geq 0$ and any $B \in \B_{\ell}$ with $|X_B| \geq k$,
$$ \left\lceil \lg \frac{1}{\mu(B)} \right\rceil -2 \leq \ell \leq  \left\lceil \lg \frac{1}{\mu(B)} \right\rceil.$$
\label{lemma:mass-vs-level}
\end{lemma}
\begin{proof}
Recall the definition of level $\ell$:
$$ B \in \B_\ell 
\ \Longleftrightarrow \ \frac{n}{2^{\ell+1}} \leq |X_B| < \frac{n}{2^\ell} 
\ \Longleftrightarrow \ 2^\ell < \frac{n}{|X_B|} \leq 2^{\ell + 1}.$$
By Lemma~\ref{lemma:ball-size-bounds}, 
$$ \frac{1}{2 \mu(B)} \leq \frac{n}{|X_B|} \leq \frac{2}{\mu(B)}.$$
Thus $2^{\ell} < 2/\mu(B)$ and $1/(2 \mu(B)) \leq 2^{\ell+1}$. These translate into the stated bounds on $\ell$.
\end{proof}

\subsection{Bounding $L_2$ using probabilistic distance: Proof of Lemma~\ref{lemma:L2-bound}}

Let $p = \dist(x, \X^{-s(x)}) \geq 2k/n$ and let $\ell = \lceil \lg (1/p) \rceil + 1 \leq \lg (2n/k)$.

For any $B \in \B_{\geq \ell}(x)$, we have $\mu(B) \leq 2|X_B|/n < 2/2^\ell \leq p$, using fidelity (Definition~\ref{defn:conformity}), the definition of sampling levels, and the definition of $\ell$, in that order. It follows that $B$ does not intersect $\X^{-s(x)}$, whereupon $s(x) \cdot \eta_X(B) \geq 0$.

Next, pick any $B \in \B_{\leq \ell}(x)$ with $s(x) \cdot \eta_X(B) < 0$; thus $B \in \B_{< \ell}(x)$. By the richness of $\B$, there exists $B' \in \B_\ell(x)$ with $X_{B'} \subset X_B$; and so $s(x) \cdot \eta_X(B') \geq 0$. 
% Thus $B$ must intersect $\X^{-s(x)}$ and has probability mass $\geq p$. We will show that there exists $B' \in \B_{\leq \ell}(x)$ such that $B' \subset B$ and $B'$ does not intersect $\X^{-s(x)}$; whereupon $s(x) \cdot \eta_X(B') \geq 0$. Indeed, take $B' \in \B(x)$ to be a subset of $B$ of $\mu$-mass $p-\epsilon$ for some very small $\epsilon$; we can do this because of the absolute continuity of $\mu$. Then $B'$ does not touch $\X^{-s(x)}$ and by Lemma~\ref{lemma:mass-vs-level}, for small enough $\epsilon$, it lies at level $\leq \ell$.

\subsection{Uncertainty region in the distributional setting: Proof of Lemma~\ref{lemma:delta-continuous}}

Recall from (\ref{eq:sampling-region}) that
$$ \Delta_\ell = \bigcup_{x \in X: L_2(x) \geq \ell} \bigcup_{B \in \B_{\ell}(x)} X_B .$$
Consider any $z \in \Delta_\ell$. Then there exists $x \in \X$ with $L_2(x) \geq \ell$ and $B \in \B_\ell(x)$ such that $z \in B$. Now, $B \in \B_\ell(x)$ implies $|X_B|/n < 1/2^\ell$ and so (by the fidelity property) $\mu(B) < 2/2^\ell$. We will show that $x \in \partial_{4/2^\ell}$ and thus $z \in \partial_{4/2^\ell, 2/2^{\ell}}$.

There are two cases for $x$. If $\dist(x, \X^{-s(x)}) < 2k/n$ then we immediately have $x \in \partial_{4/2^\ell}$ since $4/2^\ell \geq 2k/n$.

On the other hand, if $\dist(x, \X^{-s(x)}) \geq 2k/n$ then, since $L_2(x) \geq \ell$, we can apply Lemma~\ref{lemma:L2-bound} to get 
$$ \ell 
\ \leq \ \left\lceil \lg \frac{1}{\dist(x,\X^{-s(x)})} \right\rceil + 1 
\ \leq \ 
\lg \frac{1}{\dist(x,\X^{-s(x)})} + 2 .
$$
Thus $\dist(x, \X^{-s(x)}) \leq 1/2^{\ell-2}$ and $x \in \partial_{4/2^\ell}$.

\subsection{A generic label complexity bound in the distributional setting}

We will need to relate the size of the query region $\Delta_\ell$ to the probability mass of the corresponding second-order boundary. For this, we provide an analog of Lemma~\ref{lemma:level-distribution} for the distributional setting.
\begin{lemma}
With probability at least $1-2(\lg (n/k))e^{-k/3}$, the following hold for all levels $0 \leq \ell \leq \lg (n/4k)$.
\begin{enumerate}
\item[(a)] $|\{x \in X: T_x \leq \tau_\ell\}| \leq 2 n \tau_\ell$.
\item[(b)] $|\{x \in \Delta_{\ell}: T_x \leq \tau_\ell\}| \leq 2 n \mu(\partial_{4/2^\ell, 2/2^\ell}) \tau_\ell$.
\end{enumerate}
\label{lemma:level-distribution-dist}
\end{lemma}
\begin{proof} 
Part (a) is as in Lemma~\ref{lemma:level-distribution}.

For (b), pick any subset $S \subset \X$. If $X$ consists of $n$ independent draws from $\mu$, then $|\{x \in S: T_x \leq \tau_\ell\}|$ has a $\mbox{binomial}(n, \mu(S) \tau_\ell)$ distribution with expectation $n \mu(S) \tau_\ell$. The probability that it is more than twice its expected value is, by a multiplicative Chernoff bound, at most $e^{-n \mu(S) \tau_\ell/3}$. We will apply this to the various sets $S = \partial_{4/2^\ell, 2/2^\ell}$ and take a union bound over them. In each application, we will also see that $n \mu(S) \tau_\ell \geq k$.

Pick any $\ell \leq \lg (n/4k)$. If $\partial_{4/2^\ell, 2/2^\ell} = \emptyset$, then the statement in (b) is trivially true given Lemma~\ref{lemma:delta-continuous}. So assume this is not the case. Writing $p = 4/2^\ell$, we need to check that $\mu(\partial_{p, p/2}) \tau_\ell \geq k/n$, or equivalently, $\mu(\partial_{p, p/2}) \geq \max(1/2^{\ell+2}, k/n) = 1/2^{\ell+2}$. Now, $\partial_{p,p/2} \neq \emptyset \Longrightarrow \partial_p \neq \emptyset$. Pick any $x \in \partial_p$. We have by the richness of $\B$ (Definition~\ref{defn:conformity}) that there exists $B \in \B_\ell(x)$. By the definition of level, $p/8 \leq |X_B|/n < p/4$ and thus, by the fidelity property, $p/16 \leq \mu(B) < p/2$. It follows that $B \subset \partial_{p,p/2}$, whereupon and $\mu(\partial_{p,p/2}) \geq \mu(B) \geq p/16 = 1/2^{\ell+2}$.
%By absolute continuity of $\mu$, we can grow a ball around $x$ of probability mass arbitrarily close to $p/2$, so that this ball is contained within $\partial_{p,p/2}$. Thus $\mu(\partial_{p,p/2}) \geq p/2-\epsilon$ for any $\epsilon > 0$. The required conditions then follow from the value of $p$.
\end{proof}

\begin{thm}
Suppose that $k \geq (192/\gamma^2) \ln (4 |\B|/\delta)$ and that the active learning algorithm makes $0 < m \leq n$ queries. Then with probability at least $1-3\delta$, all points $x \in X$ with $L_1(x) \leq \ell_1$ and $L_2(x) \leq \ell_2$ will get Bayes-optimal labels $\yh(x) = g^*(x)$, where 
$$ \ell_1 = \left\lfloor \lg \frac{m}{32k} \right\rfloor $$
and $\ell_2$ is the largest integer $\leq \lg (n/4k)$ such that
$$ \sum_{\ell = \ell_1 + 1}^{\ell_2} 2^\ell \mu(\partial_{4/2^\ell, 2/2^\ell}) \ < \ \frac{m}{32k} .$$
\label{thm:label-complexity-dist}
\end{thm}

\begin{proof}
The proof is identical to that of Theorem~\ref{thm:label-complexity-0}; the only change is to use Lemma~\ref{lemma:level-distribution-dist}(b) in place of Lemma~\ref{lemma:level-distribution}(b).
\end{proof}

\subsection{Rates of convergence under distributional conditions}

We now give label complexity bounds under three assumptions. Let $\X_\gamma = \{x \in \X: |\eta(x)| \geq \gamma\}$; thus $X \cap \X_\gamma$ is the set of points whose assigned labels will be judged. We further divide this set by label: for $s \in \{-1,+1\}$, let $\X^s_\gamma = \{x \in \X: s \cdot \eta(x) \geq \gamma\}$.

The first assumption is on the curvature of the decision regions. It allows us to bound $L_1(x)$.
\begin{enumerate}
\item[(A1)] There is an absolute constant $p_o > 0$ for which the following holds: for any $x \in \X_\gamma$, there exists $B \in \B(x)$ such that $B \subset \X^{s(x)}_\gamma$ and $\mu(B) \geq p_o$.
\end{enumerate}
\begin{lemma}
Under (A1), every $x \in \X_\gamma$ has $L_1(x) \leq \lceil \lg (1/p_o) \rceil$, provided $n > k/2p_o$.
\label{lemma:L1-bound}
\end{lemma} 
\begin{proof}
Pick any $x \in X_\gamma$; apply (A1) to get $B \in \B(x)$ for which $\mu(B) \geq p_o$ and $B \subset \X^{s(x)}_\gamma$. By Lemma~\ref{lemma:mass-vs-level}, $B \in \B_\ell$ for $\ell \leq \lceil \lg (1/p_o) \rceil$. Now, $s(x) \cdot \eta_X(B) \geq \gamma$; moreover, for any $B' \in \B(x)$ with $X_{B'} \subset X_B$ we have $X_{B'} \subset \X^{s(x)}_\gamma$ and thus $s(x) \cdot \eta_X(B') \geq \gamma$ as well.
\end{proof}

The second assumption is a variant of the Tsybakov margin condition and bounds $|\Delta_\ell|$.
\begin{enumerate}
\item[(A2)] There are absolute constants $C > 0$ and $0 < \sigma < 1$ for which the following holds: for any $p > 0$, we have $\mu(\partial_{p,p}) \leq C p^\sigma$.
\end{enumerate}
If, for instance, $\mu$ were the uniform distribution over $[0,1]^d$, we would expect $\sigma \sim 1/d$.

The third assumption bounds the fraction of points with large $L_2$ values. It asks, what fraction of the $p$-boundary touches $\X_\gamma$? We consider two options: under (A3), the fraction is zero for $p$ small enough; under (A3'), the fraction is $p^\xi$.
\begin{enumerate}
\item[(A3)] There is an absolute constant $p_1 > 0$ such that $\mu(\partial_p \cap \X_\gamma) = 0$ for any $p < p_1$.
\item[(A3')] There are constants $C' > 0$ and $0 < \xi < 1$ such that $\mu(\partial_{p} \cap \X_\gamma) \leq C' p^\xi$ for any $p > 0$.
\end{enumerate}

Under these assumptions, we get rates of convergence as follows.
\begin{thm}
Assume (A1), (A2) and either (A3) or (A3'). There is an absolute constant $C''$ for which the following holds. Pick $0 < \delta < 1$ and take $k = O(((d \log n) + \log (1/\delta))/\gamma^2)$. If the algorithm of Fig.~\ref{alg:main} makes $m$ queries, for
$$ \frac{128k}{p_o} \ \leq \ m \ \leq \ \frac{64C \cdot 8^\sigma}{1-\sigma} \, n^{1-\sigma} k^\sigma,$$
then with probability at least $1-\delta$:
\begin{itemize}
\item Under (A3), all of $X \cap \X_\gamma$ get Bayes-optimal labels for $m > (512C/(1-\sigma)) \cdot (1/p_1)^{1-\sigma} \cdot k$.
\item Under (A3'),  Bayes-optimal labels get assigned to all but $ C'' (k/m)^{\xi/(1-\sigma)} + (2/n) \log (4/\delta)$
%$$ C'' \left(\frac{k}{m}\right)^{\xi/(1-\sigma)} + \frac{2 \log (4/\delta)}{n} $$
fraction of $X \cap \X_\gamma$.
\end{itemize}
\label{thm:label-complexity-specific}
\end{thm}
This theorem yields the result of Section~\ref{sec:Massart}. Another example is in Section~\ref{sec:Tsybakov}, where bounds are given under smoothness and Tsybakov margin conditions.

\begin{proof}
In this case, $\B$ is infinite, but by standard VC-dimension arguments there are only $O(n^{d+1})$ balls with distinct sets $X_B$. This governs the setting of $k$.

First, define $\ell_1 = \lfloor \lg (m/32k) \rfloor$ and observe that by Lemma~\ref{lemma:L1-bound}, all points in $\X_\gamma$ have 
$$ L_1(x) 
\ \leq \ 
\left\lceil \lg \frac{1}{p_o} \right\rceil
\ \leq \ 
\lg \frac{2}{p_o} 
\ \leq \ 
\lg \frac{m}{64k}
\ \leq \ 
\ell_1 .
$$
Next, pick  
$$ \ell_2 = \left\lfloor \frac{1}{1-\sigma} \left( \lg \frac{m}{32k} + \lg \frac{1-\sigma}{C \cdot 4^{1+\sigma}} \right) \right\rfloor.$$
The upper bound on $m$ ensures that this is at most $\lg (n/2k)$. Then, using (A2),
$$
\sum_{\ell = \ell_1+1}^{\ell_2} 2^\ell \mu(\partial_{4/2^\ell, 2/2^\ell})
\ \leq \ 
C \sum_{\ell = \ell_1+1}^{\ell_2} 2^\ell \left( \frac{4}{2^\ell} \right)^\sigma 
\ \leq \ 
C \cdot 4^\sigma \cdot \frac{2}{2^{1-\sigma}-1} \cdot 2^{(1-\sigma)\ell_2}
\ \leq \ 
\frac{m}{32k} .
$$
Applying Theorem~\ref{thm:label-complexity-dist}, we then see that all points in $\X_\gamma$ with $L_2(x) \leq \ell_2$ will be correctly classified. Any remaining point $x \in \X_\gamma$ has $L_2(x) > \ell_2$ and thus (by Lemma~\ref{lemma:L2-bound})
$$ \left\lceil \lg \frac{1}{\dist(x, \X^{-s(x)})} \right\rceil + 1 > \ell_2 
\ \Longrightarrow \ 
\dist(x, \X^{-s(x)}) < \frac{4}{2^{\ell_2}} \leq  \left( \frac{512C}{1-\sigma} \right)^{1/(1-\sigma)}  \cdot \left( \frac{k}{m} \right)^{1/(1-\sigma)} . 
$$
Call this quantity $p$; thus any such $x$ lies in $\partial_p$.

Under (A3), $\partial_p \cap \X_\gamma$ has zero probability mass for $p < p_1$, that is, if 
$$ \left( \frac{512C}{1-\sigma} \right)^{1/(1-\sigma)}  \cdot \left( \frac{k}{m} \right)^{1/(1-\sigma)} < p_1
\ \Longleftrightarrow \ 
m > \frac{512C}{1-\sigma} \cdot \frac{1}{p_1^{1-\sigma}} \cdot k .
$$

Under (A3'), $\mu(\partial_p \cap \X_\gamma) \leq C' p^{\xi}$; we can then apply a Bernstein bound to assert that with probability at least $1-\delta$, 
$$ |X \cap (\partial_p \cap \X_\gamma)| \ \leq \ \frac{3}{2} C' n p^{\xi} + 2 \log \frac{1}{\delta},$$
from which the bound in the theorem follows by defining $C''$ appropriately.
\end{proof}

\subsection{Label complexity under curvature and Massart noise: Proof of Theorem~\ref{thm:massart-dist}}

Let (C1) denote the strong density condition and (C2) the Massart noise and boundary condition. Theorem~\ref{thm:massart-dist} follows immediately from the more general result of Theorem~\ref{thm:label-complexity-specific}, once we establish how (C1) and (C2) relate to assumptions (A1), (A2), and (A3').

\begin{lemma}
Conditions (C1) and (C2) yield assumption (A1), with $p_o = r_o^d \cdot c_o \cdot v_d$, where $v_d$ is the volume of the unit ball in $\R^d$.
\label{lemma:A1}
\end{lemma}

\begin{proof}
Pick any $x \in \X_\gamma^+$ (the negative case is similar), and let $r = \inf_{z \in \X^0} \|x - z\|$. If $r \geq r_o$, then $B = B(x,r_o)$ is entirely in $\X_\gamma^+$. Otherwise, the reach condition (C2) implies the existence of a ball $B \subset \X_\gamma^+$ that contains $x$ and has radius $r_o$. Either way, $\mu(B) \geq c_o \vol(B) = c_o v_d r_o^d$ by (C1). 
\end{proof}

\begin{lemma}
Conditions (C1) and (C2) yield assumptions (A2) and (A3') with $\sigma = \xi = 1/d$. 
\end{lemma}

\begin{proof}
Under (C1), any ball of probability mass $\leq p$ has volume $\leq p/c_o$ and radius $\leq (p/(c_o v_d))^{1/d}$. Let's call this latter quantity $r$. Thus, any point in $\partial_p$ lies within distance $2r$ of the boundary, while a point in $\partial_{p,p}$ lies within distance $4r$.

To bound the volume of $\partial_{p,p}$, we can associate each point in this region with its nearest neighbor in $\X^0$; by condition (C1), this projection map is uniquely defined for $r < r_o/4$. The volume of the region is thus $O(r)$ and under (C1), has probability mass $O(r) = O(p^{1/d})$.
\end{proof}

\subsection{Label complexity under smoothness and Tsybakov noise}
\label{sec:Tsybakov}

Now we specialize Theorem~\ref{thm:label-complexity-specific} to another situation, where the data distribution satisfies Holder smoothness and Tsybakov margin conditions.

Recall that $\dist(x,S)$ denotes the probability-distance from $x$ and set $S$. We will overload notation so that for $x,x' \in \X$,
$$ \dist(x,x') = \dist(x, \{x'\}) = \inf\{\mu(B): B \in \B(x) \cap \B(x')\},$$
that is, the probability mass of the smallest ball containing both $x$ and $x'$. We will impose the following conditions on the data.
\begin{enumerate}
\item[(C1)] [Strong density condition] This was introduced in the earlier (Massart) example.
\item[(C2')] [Holder-smoothness of conditional probability function] There exist constants $L, \alpha$ such that
$$ |\eta(x) - \eta(x')| \ \leq \ L \cdot \dist(x,x')^\alpha$$
for all $x,x' \in \X$.
\item[(C3')] [Tsybakov margin condition] There exists constants $M, \beta$ such that
$$ \mu(\{x \in \X: |\eta(x)| \leq \tau\}) \leq M \tau^\beta$$
for all $\tau \in (0,1)$.
\item[(C4')] [Bounded curvature] The boundaries $\{x \in \X: \eta(x) = \gamma\}$ and $\{x \in \X: \eta(x) = -\gamma\}$ are $(d-1)$-dimensional Riemannian manifolds of reach $r_o > 0$.
\end{enumerate}

Lemma~\ref{lemma:A1} continues to hold, with condition (C4') doing the job of (C2). This yields assumption (A1). For the remaining assumptions, we first obtain a consequence of the Holder condition.

\begin{lemma}
Under (C2'), for any $p, q > 0$,
\begin{enumerate}
\item[(a)] $x \in \partial_p \implies |\eta(x)| \leq L p^\alpha$.
\item[(b)] $x \in \partial_{p,q} \implies |\eta(x)| \leq L(p^\alpha + q^\alpha)$.
\end{enumerate}
\label{lemma:holder-consequence}
\end{lemma}
\begin{proof}
Pick any $x \in \partial_p$. Let $s(x) = \mbox{sign}(\eta(x))$. By definition of the $p$-boundary set, for any $\epsilon > 0$, there exists $x' \in \X^{-s(x)}$ such that $\dist(x,x') < p+\epsilon$. By the Holder condition, $|\eta(x) - \eta(x')| < L(p+\epsilon)^\alpha$ and thus $|\eta(x)| < L(p+\epsilon)^\alpha$. Since this holds for any $\epsilon > 0$, we get part (a).

For (b), pick $x \in \partial_{p,q}$ and $\epsilon > 0$. Then there exists $x' \in \partial_p$ with $\dist(x,x') < q+\epsilon$. As before, we use the Holder condition to conclude that $|\eta(x)| < |\eta(x')| + L(q+\epsilon)^\alpha$ and then invoke (a).
\end{proof}

\begin{lemma}
Conditions (C2') and (C3') yield assumptions (A2) and (A3) with $C = (2L)^\beta M$, $\sigma = \alpha \beta$, and $p_1 = (\gamma/L)^{1/\alpha}$.
\label{lemma:boundary-constants} 
\end{lemma}

\begin{proof}
By Lemma~\ref{lemma:holder-consequence}, $\partial_{p,p} \subset \{x \in \X: |\eta(x)| \leq 2Lp^\alpha\}$; the probability mass of this set can be bounded by (C3').

Also by Lemma~\ref{lemma:holder-consequence}, $\partial_p$ is entirely contained in $\{x \in \X: |\eta(x)| \leq L p^\alpha\}$. For $p < p_1$, this does not intersect $\X_\gamma$.
\end{proof}

\begin{thm}
Assume conditions (C1) and (C2')--(C4'). There are constants $c_2, c_3$ for which the following holds. Pick $0 < \delta < 1$ and take $k = O(((d \log n) + \log (1/\delta))/\gamma^2)$. Suppose the algorithm of Fig.~\ref{alg:main} makes $m$ queries, where
$$ \frac{c_2k}{p_1^{1-\sigma}} \ \leq \ m \ \leq \ c_3 n^{1-\sigma} ,$$
with $p_1, \sigma$ as given in Lemma~\ref{lemma:boundary-constants}. Then with probability at least $1-\delta$, all of $X \cap \X_\gamma$ is assigned Bayes-optimal labels.
\label{thm:smoothness-dist}
\end{thm}

In this setting, regular $k$-nearest neighbor classification would also correctly classify all of $X \cap X_\gamma$ if given $O(k/p_1)$ random labeled points. The benefit of active learning is to thus reduce the label complexity to $O(k/p_1^{1-\alpha\beta})$.
\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: